\documentclass[12pt,letter]{letter}

\include{setup}

\signature{\vspace{-40pt} Gaurav Vaidya, Ph.D.}
%\address{3611 SW 34th St Apt 72, \\ Gainesville FL 32608}

\begin{document}

\begin{letter}{Jordan Todd, HR Consultant, \\
Vice Chancellor for Research UBC Human Resources, \\
University of North Carolina, 
Chapel Hill NC}

\opening{Dear Mr. Todd:}

Thank you so much for this opportunity to apply for the position of Semantics Web Technologist at the Renaissance Computing Institute (RENCI). I was extremely excited to see this job advertised, as I have considerable experience with both scientific software development and the semantic web and have been looking for opportunities to work in both of these areas outside of a traditional academic lab. I love how much emphasis RENCI places on scientific software development in its 2017 Strategic Plan, and I think that working at RENCI will help me learn from a community of scientific software developers and will allow me apply my skills across a variety of disciplines.

I have had a number of work experiences that make me particularly well-suited for this position. I worked on data modeling throughout my Ph.D., and developed an XML schema to build, store and analyse the datasets I built for my dissertation. In summer 2012, I worked with librarians at the Missouri Botanical Gardens on its Art of Life project to improve the metadata available on Biodiversity Heritage Library images on Flickr and the Wikimedia Commons, an online repository of millions of images, audio and video files that are reusable under Creative Commons licenses. In summer 2014, I worked on a project at DBpedia, an online repository of facts extracted from Wikipedia and stored in a triple-store in the Resource Description Framework (RDF) format and accessible using a SPARQL interface. My project expanded DBpedia's extractors, written in the Scala programming language, to be able to extract facts stored in the Wikimedia Commons. For the last year and a half, I have worked full-time as a post-doctoral associate and lead software developer on the Phyloreferencing project, whose goal is to develop an ontology of computable, testable and semantically-rich definitions for groups of related living organisms called `clades', which we plan to publish in the Web Ontology Language (OWL). I am building the software tools needed for this project using JavaScript on Node.js, Vue CLI, Java and Python. Please see my attached curriculum vitae and references for more details and other relevant experience.

I hope that my interests in teaching and in open source projects such as Wikipedia will also be of use to RENCI. I was very excited to see how much emphasis RENCI placed on making scientific data widely accessible, such as through your Galapagos Science Center and your Emerging Technologies for Higher Education white paper, and I would love to help with further outreach activities.

I should note that as an international student in the United States currently working under the F-1 OPT program, I will require sponsorship for an H-1 visa if chosen for this position.

Thank you so much for your attention!

\closing{Yours sincerely,}

\end{letter}

\end{document}